---
title: "Calculate aMAP Score"
author: "JC"
date: "2023-10-17"
output: html_document
---
```{r}
source("../config.R") #Loads libraries, variables and global function
```



# Importing relevant columns

```{r}
setwd(sharepoint_ukb) #from central storage
df_withdrawals <- read.delim("raw/withdrawals.txt")
df_blood <- fread("extracted/UKB_Patients_blood.csv") %>% check_and_remove_withdrawals(df_withdrawals)
df_afp <- fread("extracted/AFP.csv") %>% check_and_remove_withdrawals(df_withdrawals) %>%          #just for AMAP2
  select(-V1) %>%                    # Remove the "V1" column
  rename(AFP = `66`) 

setwd(project_path) #project-specific
load("data/dataframes/df_covariates.RData")
df_covariates <- check_and_remove_withdrawals(df_covariates, df_withdrawals)
load("data/dataframes/df_icd.RData")


df_y <- read.csv(paste0(project_path, "/data/dataframes/df_y.csv")) %>% check_and_remove_withdrawals(df_withdrawals) %>%
      select(c(eid, status, status_cancerreg))


df_icd <- df_icd %>%
  mutate(across(-eid, ~ as.factor(.x))) %>%
  check_and_remove_withdrawals(df_withdrawals)


df_blood <- fread(paste0(sharepoint_ukb, "/extracted/UKB_Patients_blood.csv")) %>% check_and_remove_withdrawals(df_withdrawals)
Blood_Marker_Index <- read_excel(paste0(sharepoint_ukb, "/", master_table), sheet="Blood count and biochemistry") %>%
  mutate(datafield = paste0(datafield, '-0.0'))

df_blood <- df_blood %>%
  rename_with(
    ~ Blood_Marker_Index$Description[match(., Blood_Marker_Index$datafield)],
    .cols = -eid  # Exclude 'eid' from renaming
  )

df_blood <- df_blood %>%
  select(eid, all_of(Blood_Marker_Index$Description)) %>% # Order Blood columns
  mutate_at(vars(-1), mean.impute)

```



# Extract aMAP Score
Calculate aMAP (Advanced Model for End-Stage Liver Disease Allocation) score.

  The aMAP score is a formula used to assess the severity of liver disease. It takes into account
    the patient's age, gender, bilirubin level, albumin level, and platelet count to provide a
    numerical score.

  Parameters:
  - age (int or float): Age of the patient in years.
  - gender (int): Gender of the patient (0 for female, 1 for male).
  - bilirubin (float): Bilirubin level in micromoles per liter (μmol/L).
  - albumin (float): Albumin level in grams per liter (g/L).
  - platelet_count (float): Platelet count in thousands per cubic millimeter (10^3/mm³).
  
  Formula:
    aMAP = ((age * 0.06 + gender * 0.89 + 0.48 * ((log10(bilirubin) * 0.66) + (albumin * -0.085)) - 0.01 * platelet_count) + 7.4) / 14.77 * 100

```{r}
#Access relevant columns and impute NAs
df_amap <- df_blood %>% 
  select(c("eid", "Platelet count", "Albumin", "Total bilirubin")) %>%
  rename(
    bilirubin = `Total bilirubin`,
    platelet_count = `Platelet count`,
    albumin = `Albumin`
  ) %>%
  left_join(df_covariates %>% select(c("eid", "AGE", "SEX")), by = "eid") %>% 
  rename(gender = SEX)

df_amap <- as.data.frame(df_amap)
df_amap[,2:4] <- mean.impute(df_amap[,2:4])
df_amap$age <- df_covariates$AGE
df_amap$gender <- df_covariates$SEX


df_amap$gender <- ifelse(df_amap$gender == "Female", 0, 1)
df_amap$gender <- as.numeric(df_amap$gender)



#Calculate value
df_amap$aMAP <- ((df_amap$age * 0.06 + df_amap$gender * 0.89 + 0.48 * ((log10(df_amap$bilirubin) * 0.66) + (df_amap$albumin * -0.085)) - 0.01 * df_amap$platelet_count) + 7.4) / 14.77 * 100

df_amap <- df_amap %>% 
  select(c("eid", "aMAP", "gender")) %>%
  mutate(aMAP = aMAP / 100)
  

summary(df_amap$aMAP)

```

# Subset for PAR predefined in Table X 
# Export amap dataframes
```{r}




df_amap <- inner_join(df_amap, df_y, by="eid")
write.csv(df_amap, file=paste(project_path, "/Models/df_amap.csv", sep=''))

df_amap_par <- inner_join(df_amap, par_eids, by="eid")
write.csv(df_amap_par, file=paste(project_path, "/Models/df_amap_par.csv", sep=''))


```



# Calculate Fib-4
In Sterling et al, which studied patients co-infected with HIV and Hepatitis C, a simple, non-invasive index (FIB-4) was developed to identify significant hepatic fibrosis.

The formula for FIB-4 is: Age ([yr] x AST [U/L]) / ((PLT [10(9)/L]) x (ALT [U/L])(1/2)).

The AUROC of the index was 0.765 for differentiation between Ishak stage 0-3 and 4-6. At a cutoff of <1.45 in the validation set, the negative predictive value to exclude advanced fibrosis (stage 4-6) was 90% with a sensitivity of 70%. A cutoff of >3.25 had a positive predictive value of 65% and a specificity of 97%. Using these cutoffs, 87% of the 198 patients with FIB-4 values outside 1.45-3.25 would be correctly classified, and liver biopsy could be avoided in 71% of the validation group.

https://www.qxmd.com/pubmed/16729309 Sterling et al Hepatology 2006

```{r}
df_fib4 <- df_covariates %>% select("eid", "AGE") %>%
  left_join(df_blood %>% 
              select(eid, 
                     AST = `Aspartate aminotransferase`,
                     ALT = `Alanine aminotransferase`,
                     PLT = `Platelet count`),
            by = "eid")

df_fib4 <- df_fib4 %>%
  mutate(
    FIB4_orig = (AGE * AST) / (PLT * (ALT * 0.5))
  )

# Calculate normalized FIB-4 score (0 to 1)
df_fib4 <- df_fib4 %>%
  mutate(
    FIB4 = (FIB4_orig - min(FIB4_orig, na.rm = TRUE)) / (max(FIB4_orig, na.rm = TRUE) - min(FIB4_orig, na.rm = TRUE))
  )

sum(is.na(df_fib4$FIB4))
summary(df_fib4$FIB4)
```

# Calculate APRI
(AST to Platelet Ratio Index):

Formula: (AST / ULN_AST) / (PLT / 100)

(Sex-specific ULNs)
```{r}
df_apri <- df_covariates %>% 
  select(eid, SEX) %>%
  left_join(df_blood %>% 
              select(eid, 
                     AST = `Aspartate aminotransferase`,
                     PLT = `Platelet count`),
            by = "eid")

df_apri <- df_apri %>%
  mutate(
    ULN_AST = case_when(
      SEX == "Female" ~ 35,
      SEX == "Male" ~ 50,
      TRUE ~ NA_real_
    ),
    APRI_orig = (AST / ULN_AST) / (PLT / 100)
  )

# Normalize APRI score (0 to 1)
df_apri <- df_apri %>%
  mutate(
    APRI = (APRI_orig - min(APRI_orig, na.rm = TRUE)) / (max(APRI_orig, na.rm = TRUE) - min(APRI_orig, na.rm = TRUE))
  )

# Summary of APRI scores
summary(df_apri)
```

NFS Score:
AST, ALT, age, BMI,
T2DM, platelets,
albumin

NAFLD fibrosis score = -1.675 + 0.037 × age (year) + 0.094 × BMI (kg/m2) + 1.13 × IFG/diabetes (yes = 1, no = 0) + 0.99 × AST/ALT ratio - 0.013 × platelet count (×109/L) - 0.66 × albumin (g/dL)

Interpretation:

<-1.455: Low probability of advanced fibrosis
-1.455 to 0.676: Indeterminate score
0.676: High probability of advanced fibrosis

```{r}
df_nfs <- df_covariates %>% 
  select(eid, SEX, BMI, AGE) %>%
  left_join(df_icd %>% select(eid, DM,)) %>%
  left_join(df_blood %>% 
              select(eid, 
                     AST = `Aspartate aminotransferase`,
                     ALT = `Alanine aminotransferase`,
                     PLT = `Platelet count`,
                    Albumin),
            by = "eid")

df_nfs <- df_nfs %>%
  mutate(
    DM_numeric = as.numeric(DM),
    AST_ALT_ratio = AST / ALT,
    NFS_orig = -1.675 + 0.037 * AGE + 0.094 * BMI + 1.13 * ifelse(DM_numeric == 1, 1, 0) + 
          0.99 * AST_ALT_ratio - 0.013 * PLT - 0.66 * Albumin
  )

# Normalize NFS score (0 to 1)
df_nfs <- df_nfs %>%
  mutate(
    NFS = (NFS_orig - min(NFS_orig, na.rm = TRUE)) / (max(NFS_orig, na.rm = TRUE) - min(NFS_orig, na.rm = TRUE))
  )

# Summary of NFS scores
summary(df_nfs$NFS)
```

# Merge scores
```{r}
df_benchmark <- df_amap %>%
  select(eid, aMAP, gender) %>%
  inner_join(df_apri %>% select(eid, APRI), by = "eid") %>%
  inner_join(df_fib4 %>% select(eid, FIB4), by = "eid") %>%
  inner_join(df_nfs %>% select(eid, NFS), by = "eid") %>%
  inner_join(df_icd %>% select(eid, `Liver cirrhosis`), by = "eid")

df_benchmark <- merge(df_benchmark, df_y, by="eid")

# Summary of the benchmark dataframe
summary(df_benchmark)

# Correlation matrix of the normalized scores
cor_matrix <- cor(df_benchmark[, -1], use = "complete.obs")
print(cor_matrix)
```










# Subset for liver cirrhosis/NAFLD
```{r}
# subset= "Other and unspecified cirrhosis of liver"
# df_merged <- merge(df_amap, df_icd[, c("eid", "Other and unspecified cirrhosis of liver")], by="eid", all.x = TRUE)
# df_amap_cirrhosis <- df_merged[df_merged$'Other and unspecified cirrhosis of liver' == 1, ]
# df_amap <- merge(df_amap, df_y, by="eid")
# 
# df_merged <- merge(df_amap, df_icd[, c("eid", "NAFLD")], by="eid", all.x = TRUE)
# df_amap_nafld <- df_merged[df_merged$'NAFLD' == 1, ]
# df_amap <- merge(df_amap, df_y, by="eid")
#   
# write.csv(df_amap_cirrhosis, file=paste(project_path, "/data/dataframes/df_amap_cirrhosis.csv", sep=''))
# write.csv(df_amap_nafld, file=paste(project_path, "/data/dataframes/df_amap_nafld.csv", sep=''))

```



Export
```{r}
write.csv(df_amap, file=paste(project_path, "/data/dataframes/df_amap.csv", sep=''))

write.csv(df_benchmark, file=paste(project_path, "/data/dataframes/df_benchmark.csv", sep=''))



par_eids <- read.csv(file.path(project_path, "data/07_05_2024/par_eids.csv")) %>%
  rename("eid" = "x")
df_amap_par <- inner_join(df_amap, par_eids, by="eid")
write.csv(df_amap_par, file=paste(project_path, "/data/dataframes/df_amap_par.csv", sep=''))


```

