---
title: "Extract multiple diagnosis"
author: "JC"
date: "2024-05-04"
output: html_document
---

```{r}
source("../config.R") #Loads libraries, variables and global function

```



# Load HESIN
```{r}
#oeffne das Datenset mit Diagnosen der Patienten, hesin_diag. Diese Datei wird ständig aktualisiert, und ist nicht besonders groß, daher am besten immer neu aus raw-Datei.
setwd(sharepoint_ukb)
df_withdrawals <- read.delim(withdrawals_path)

cols_to_read <- c("eid", "ins_index", "diag_icd10", "diag_icd9") #predefine cols for hesin/hesin_diag to be read
hesin_diag <- fread(hesin_diag_path, select = cols_to_read) %>% check_and_remove_withdrawals(df_withdrawals)

cols_to_read <- c("eid", "ins_index", "epistart")
hesin <- fread(hesin_path, select = cols_to_read) %>% check_and_remove_withdrawals(df_withdrawals)
setkey(hesin_diag, eid, ins_index)
setkey(hesin, eid, ins_index)
pat_icds <- merge(hesin_diag, hesin, by = c("eid", "ins_index"), all = TRUE)


rm(hesin, hesin_diag)
```


# Transform HESIN
```{r}
# setwd(sharepoint_ukb)
# df_basic <- fread("extracted/basic_info.csv") %>% select(c("eid", "53-0.0")) %>% check_and_remove_withdrawals(df_withdrawals) %>%
#   rename(assessment = `53-0.0`) %>%  # Change column name of time  
#   mutate(assessment = as.Date(assessment, format = "%Y-%m-%d"))

setwd(sharepoint_ukb)
df_basic <- fread("extracted/basic_info.csv") %>% check_and_remove_withdrawals(df_withdrawals) %>%
  mutate(assessment = as.Date(assessment, format = "%Y-%m-%d"))
  


pat_icds <- merge(pat_icds, df_basic, by = "eid", all.x = TRUE) # merge with pat_icds dataframe
rm(df_basic)
pat_icds[, epistart := as.IDate(epistart, format = "%d/%m/%Y")]
pat_icds[, difftime := as.numeric(difftime(assessment, epistart, units = "weeks"))/52.25]

pat_icds_all <- pat_icds #stores all info here, as pat_icds itself will be cut-off at a defined timepoint. 

pat_icds <- pat_icds[difftime >= -(timeframe)] #subsets our ICD info to only obtain data from the first {timeframe} years after initial assessment, e.g. 5 years, starting at baseline
#pat_icds <- pat_icds[, .SD, .SDcols = !c("difftime", "assessment")] #remove unnecessary columns

```

# Load Metadata (: Create an excel sheet for all the diagnosis that you want to look into, before continuing here)
```{r}
#oeffne Projektspezifische ICD Schlüssel Datasets
setwd(sharepoint_ukb)
ICD_Groups <- read_excel(icd_dict_path, sheet = "ICD_Groups") %>% select(-c("n", "Summe"))
ICD_Singles <- read_excel(icd_dict_path, sheet = "ICD_Singles") %>% select(-c("n", "Summe"))
Patients_at_risk <- read_excel(icd_dict_path, sheet = "Patients at risk")


par_subset <- par_subset[par_subset!="Blood_Parameters"] #Remove the Blood parameters as diagnosis label, only hard-coded diagnosis are of itnerest here

Patients_at_risk <- Patients_at_risk[Patients_at_risk$Group %in% par_subset, ]
par_icd_codes <- Patients_at_risk$ICD10

```

#ICD Starts with (wenn es eine ganze Gruppe von ICD10 codes ist dann diese code, die Anforderungen koennen beliebig lang hintereinander ergaenzt werden je nachdem wie viele Gruppen der ICD code hat)
```{r}
icd_dm<-subset(pat_icds, startsWith(as.character(pat_icds$diag_icd9), '250')|startsWith(as.character(pat_icds$diag_icd10), 'E10')| startsWith(as.character(pat_icds$diag_icd10), 'E11')| startsWith(as.character(pat_icds$diag_icd10), 'E12')| startsWith(as.character(pat_icds$diag_icd10), 'E13') | startsWith(as.character(pat_icds$diag_icd10), 'E14'))
          
icd_dm<- as.data.frame(icd_dm[!duplicated(icd_dm$eid),]) # #Step 2: Duplikate entfernen: 
icd_dm$DM <- 1

icd_dm<-select(icd_dm, c(eid, DM))
```

# ICD Groups (e.g. many icds for all subtypes of gastric cancer shall be fused in information "Gastric Cancer y/n)
```{r}
# Merge diagnosis of interest with all_pat_all_diag df (hesin_diag). 
#Left_join auf Diagnosen -> alle nicht erwähnten Diagnosen "fallen raus", alle Duplikate werden entfernt
pat_icds_labelled <- left_join(x= ICD_Groups, y= pat_icds, by = c("ICD10" = "diag_icd10")) %>% distinct()

#Check for amount of diagnosis per actual ICD, instead of group. Likely here will be the correct amounts, while summarized will be less due to parallel coded diagnosis
df_icd_groups_before <- pat_icds_labelled %>% group_by(Diagnosis, eid) %>% summarise(occurence = n()) %>% spread(Diagnosis, occurence) 
sum_groups_before <- as.data.frame(colSums(df_icd_groups_before, na.rm=TRUE))

# #groupby + summarises sortiert Dataset nach Diagnosegruppen und eid,  spread macht aus langem schmalen Dataset ein breites mit logical/boolean values
df_icd_groups <- pat_icds_labelled %>% group_by(Group, eid) %>% summarise(occurence = n()) %>% spread(Group, occurence) 
df_icd_groups[df_icd_groups > 1 & df_icd_groups < 100]<- 1 
df_icd_groups[is.na(df_icd_groups)] <- 0
sum_groups_after <- as.data.frame(colSums(df_icd_groups))
rm(pat_icds_labelled, df_icd_groups_2)
```
# ICD Singles (if requested entity is diagnosis itself (e.g. NASH, NAFLD etc, or every subgroup should still behave as subgroup)
```{r}
# Merge diagnosis of interest with all_pat_all_diag df (hesin_diag). 
#Left_join auf Diagnosen -> alle nicht erwähnten Diagnosen "fallen raus", alle Duplikate werden entfernt
pat_icds_labelled <- left_join(x= ICD_Singles, y= pat_icds, by = c("ICD10" = "diag_icd10")) %>%  distinct() 

# #groupby + summarises sortiert Dataset nach Diagnosegruppen und eid,  spread macht aus langem schmalen Dataset ein breites mit logical/boolean values
df_icd_singles <- pat_icds_labelled %>% group_by(Diagnosis, eid) %>% summarise(occurence = n()) %>% spread(Diagnosis, occurence) 
df_icd_singles[df_icd_singles > 1 & df_icd_singles < 10000]<- 1 
df_icd_singles[is.na(df_icd_singles)] <- 0
sum_singles <- as.data.frame(colSums(df_icd_singles))
rm(pat_icds_labelled)
```

#Merging groups and singles
```{r}
df_icd <- full_join(df_icd_singles, df_icd_groups, by = "eid")
df_icd <- merge(df_icd, icd_dm, by="eid", all=TRUE)
df_icd <- replace(df_icd, is.na(df_icd), 0)
```

#Questionnaire info on diagnosis (e.g. "diabetes diagnosed by doctor")
```{r}
setwd(sharepoint_ukb)
covariates <- fread("extracted/covariates.csv")
df_dmdoc <- dplyr::select(covariates, eid, `2443-0.0`) %>%
  rename(`DM Questionnaire` = `2443-0.0`) %>%
  mutate(`DM Questionnaire` = if_else(`DM Questionnaire` == 1, 1, 0))

pre_merge_cases <- sum(df_icd$DM == 1) # Count positive cases before merge
print(paste("Number of positive cases of 'DM' before merge:", pre_merge_cases))
df_icd <- merge(df_icd, df_dmdoc, by="eid", all.x=TRUE)
df_icd$DM <- pmax(df_icd$DM, df_icd$`DM Questionnaire`, na.rm = TRUE) # Update the 'DM' column
post_merge_cases <- sum(df_icd$DM == 1, na.rm = TRUE)
print(paste("Number of positive cases of 'DM' after merge:", post_merge_cases)) # Count positive cases after merge
df_icd <- df_icd %>%
  select(-"DM Questionnaire")
```

# Merge NHS data onto eid data
(All patients agreed to have NHS data transferred to UKB, so if not in NHS -> No diagnosis)
```{r}
setwd(sharepoint_ukb)
df_eid <- fread("extracted/UKB_Patients_eids.csv") %>% check_and_remove_withdrawals(df_withdrawals)
df_icd <- merge(df_eid, df_icd, by="eid", all.x =TRUE)
print(paste("Nr of patients without any NHS entries:", nrow(df_icd) - sum(complete.cases(df_icd))))
df_icd[is.na(df_icd)] <- 0
df_icd <- df_icd[!duplicated(df_icd), ]

# Compress values > 1 to 1
for (column in names(df_icd)){
  if (column != "eid" && is.numeric(df_icd[[column]])){
    df_icd[[column]][df_icd[[column]] > 1] <- 1
  }
}
```
# Self-reported diagnosis
```{r}
# Create index table
# self_diag_index <- fread(file.path(sharepoint_ukb, "meta/coding_selfreported_diag.tsv"))
# 
# 
# self_diag_categories <- self_diag_index %>% filter(coding==-1)
# 
# 
# # Create column "category" for overview, out of tree-based structure
# self_diag_index <- self_diag_index %>% left_join(self_diag_categories, by = c("parent_id" = "node_id"), suffix = c("", ".y")) %>% select(-coding.y, -parent_id.y, -selectable.y) %>%
#   rename(category= meaning.y)
# 
# update_category <- function(df) {
#   for (i in 1:nrow(df)) {
#     if (is.na(df$category[i])) {
#       # Look up parent's category
#       parent_category <- df$category[df$node_id == df$parent_id[i]]
#       if (length(parent_category) > 0 && !is.na(parent_category[1])) {
#         df$category[i] <- parent_category[1]
#       }
#     }
#   }
#   return(df)
# }
# 
# previous_na_count <- sum(is.na(self_diag_index$category))
# repeat {
#   self_diag_index <- update_category(self_diag_index)
#   current_na_count <- sum(is.na(self_diag_index$category))
#   if (current_na_count == previous_na_count) {
#     break
#   }
#   previous_na_count <- current_na_count
# }
# self_diag_index <- self_diag_index %>% filter(coding!= -1)
# 
# sheets_list <- list(
#   "Diagnosis Index" = self_diag_index,
#   "Categories" = self_diag_categories
# )
#write_xlsx(sheets_list, file.path(sharepoint_ukb, "meta/self_coded_diags.xlsx")) Don't overwrite this excel file please 


# Extract all self-reported diagnosis data

# setwd(sharepoint_ukb)
# column_names <- fread("raw/ukb52200.csv", nrows = 0)
# pattern <- "^20002-0\\..*$"
# matching_columns <- grep(pattern, names(column_names), value = TRUE)
# df_self_diag<- fread("raw/ukb52200.csv", select=c("eid", matching_columns)) %>% check_and_remove_withdrawals(df_withdrawals)
# save(df_self_diag, file=paste0(sharepoint_ukb, "/extracted/df_self_diag.RData"))

load(paste0(sharepoint_ukb, "/extracted/df_self_diag.RData"))
self_diag_index <- read_excel(paste0(sharepoint_ukb, "/meta/self_coded_diags.xlsx"), sheet="Diagnosis Index")

df_self_diag <- df_self_diag %>% #Create a long table with multiple rows per eid, as in hesin_diag
  pivot_longer(
    cols = starts_with("20002"),
    names_to = "diag_nr",
    values_to = "diagnosis"
  ) %>% 
  filter(!is.na(diagnosis)) #Remove empty rows
  
# Label diagnosis
df_self_diag_labelled <- left_join(x=df_self_diag, y=self_diag_index, by = c("diagnosis" = "coding"), suffix = c("", ".y")) %>%
  filter(!!sym(project_key) == 'y') %>%
  select(-parent_id, -selectable, -node_id)

# Index self-reported diagnosis on df_icd
df_self_diag_groups <- left_join(df_self_diag_labelled, ICD_Groups, by=c("ICD_Code" = "ICD10"), suffix= c("", ".y")) %>% 
  filter(!is.na(Group)) %>%
  select(c(eid, Group)) %>%
  rename(Diagnosis = Group) %>%
  mutate(value = 1) %>%
  pivot_wider(
    names_from = Diagnosis,
    values_from = value,
  ) %>%
  mutate(across(-eid, ~map_dbl(.x, ~if (is.null(.x)) 0 else 1)))

df_self_diag_singles <- left_join(df_self_diag_labelled, ICD_Singles, by=c("ICD_Code" = "ICD10"), suffix= c("", ".y")) %>% 
  filter(!is.na(Diagnosis)) %>%
  select(c(eid, Diagnosis)) %>%
  mutate(value = 1) %>%
  pivot_wider(
    names_from = Diagnosis,
    values_from = value,
  ) %>%
  mutate(across(-eid, ~map_dbl(.x, ~if (is.null(.x)) 0 else 1)))

df_self <- full_join(df_self_diag_groups, df_self_diag_singles, by = "eid") %>%
  right_join(df_eid, by= "eid") %>%
  mutate(across(-eid, ~replace_na(.x, 0))) %>%
  as.data.frame() #%>%
  #mutate(across(-eid, ~ as.factor(.x)))

check_factor_levels(df_self)

sum_self <- create_summary(df_self)

rm(df_self_diag_groups, df_self_diag_singles, df_self_diag_labelled, df_self_diag)

#setdiff(colnames(df_icd), colnames(df_self)) check the difference to df_icd

```



# Subsets and Export
```{r}
#df_diagnosis <- df_icd #only necessary when no merge of NHS and self-reported data
setwd(project_path)
load("data/dataframes/df_blood.RData")
load("data/dataframes/df_covariates.RData")


# if (project_key == "hcc"){
#     df_blood_risk <- df_blood %>%
#   select(c("eid", "Aspartate aminotransferase", "Alanine aminotransferase", "Gamma glutamyltransferase" )) %>%
#   inner_join(select(df_covariates, eid, SEX), by = "eid") %>%
#   mutate(
#     Elevated_AST = if_else(SEX == "Female" & `Aspartate aminotransferase` > 35 | SEX == "Male" & `Aspartate aminotransferase` > 50, 1, 0),
#     Elevated_ALT = if_else(SEX == "Female" & `Alanine aminotransferase` > 35 | SEX == "Male" & `Alanine aminotransferase` > 50, 1, 0),
#     Elevated_GGT = if_else(SEX == "Female" & `Gamma glutamyltransferase` > 40 | SEX == "Male" & `Gamma glutamyltransferase` > 60, 1, 0)
#   ) %>%
#   mutate(Elevated_Liver_Enzymes = if_else(Elevated_AST == 1 | Elevated_ALT == 1 | Elevated_GGT == 1, 1, 0)) %>%
#   select(-c("Aspartate aminotransferase", "Alanine aminotransferase", "Gamma glutamyltransferase", "SEX"))
# }


if (project_key == "cca") {   
  df_blood_risk <- df_blood %>%
  select(c("eid", "Aspartate aminotransferase", "Alanine aminotransferase", "Gamma glutamyltransferase", "Alkaline phosphatase" )) %>%
  inner_join(select(df_covariates, eid, SEX), by = "eid") %>%
  mutate(
    Elevated_AST = if_else(SEX == "Female" & `Aspartate aminotransferase` > 35 | SEX == "Male" & `Aspartate aminotransferase` > 50, 1, 0),
    Elevated_ALT = if_else(SEX == "Female" & `Alanine aminotransferase` > 35 | SEX == "Male" & `Alanine aminotransferase` > 50, 1, 0),
    Elevated_GGT = if_else(SEX == "Female" & `Gamma glutamyltransferase` > 40 | SEX == "Male" & `Gamma glutamyltransferase` > 60, 1, 0),
    Elevated_AP = if_else(SEX == "Female" & `Alkaline phosphatase` > 105 | SEX == "Male" & `Alkaline phosphatase` > 130, 1, 0)
  ) %>%
  mutate(Elevated_Liver_Enzymes = if_else(Elevated_AST == 1 | Elevated_ALT == 1 | Elevated_GGT == 1, 1, 0)) %>%
  select(-c("Aspartate aminotransferase", "Alanine aminotransferase", "Gamma glutamyltransferase", "SEX", "Alkaline phosphatase"))
  
}
#Append df_self data
df_diagnosis <- full_join(df_icd, df_self, by = "eid", suffix = c(".icd", ".self"))
icd_cols <- names(df_diagnosis)[str_detect(names(df_diagnosis), "\\.icd$")]
self_cols <- names(df_diagnosis)[str_detect(names(df_diagnosis), "\\.self$")]
df_diagnosis <- inner_join(df_diagnosis, df_blood_risk, by = "eid")

# Remove suffixes to identify common columns
common_base_cols <- sub("\\.icd$", "", icd_cols)
common_cols_icd <- paste0(common_base_cols, ".icd")
common_cols_self <- paste0(common_base_cols, ".self")

# Vectorized approach to coalesce .icd and .self columns
for(i in seq_along(common_cols_icd)) {
  icd_col <- common_cols_icd[i]
  self_col <- common_cols_self[i]
  
  df_diagnosis[[icd_col]] <- with(df_diagnosis, pmax(get(icd_col), get(self_col), na.rm = TRUE))
}

# Clean df
df_diagnosis <- df_diagnosis %>%
  select(-all_of(common_cols_self)) %>%
  rename_with(~str_remove(.x, "\\.icd$")) %>%
  convert_to_factor()

#Create file paths
sum_path <- paste0(project_path, "/supplement/ICD_occurences.xlsx")
df_diagnosis_path <- paste0(project_path, "/data/dataframes/UKB_Patients_ICD_All.csv")


# Create subset dataframes
df_positives <- innerjoin_df_y(df_diagnosis) #Innerjoin on positive DOI Cases
df_par <- filter_rows_with_pos_entries(df_diagnosis)
innerjoin_df_y(df_par) #check to see how many positives are in your subset

#Create summaries
sum_par <- create_summary(df_par)
sum_diagnosis <- create_summary(df_diagnosis)
sum_positives <- create_summary(df_positives)


# Write summary files
write.xlsx(sum_diagnosis, file = sum_path, sheetName = "All", row.names = FALSE)
write.xlsx(sum_positives, file = sum_path, sheetName = "Positives", append = TRUE, row.names = FALSE)
write.xlsx(sum_par, file = sum_path, sheetName = "PAR", append = TRUE, row.names = FALSE)


save(df_icd, file=paste(project_path, "/data/dataframes/df_icd.RData", sep=''))
save(df_diagnosis, file=paste(project_path, "/data/dataframes/df_diagnosis.RData", sep=''))


# Export the eids for patients at risk

par_eids <- df_par %>% select(eid) # Export table for patients at risk subset
innerjoin_df_y(par_eids)

save(par_eids, file=paste0(project_path, "/data/dataframes/df_par_", timeframe, "_years.RData"))


```




# Vis Summary of Occurence of all diagnosis
```{r}

plot_all <- barplot_diags(df_diagnosis, diags_to_remove = c("Arterial hypertension", "DM"))

plot_positives <- barplot_diags(df_positives, diags_to_remove = c("Arterial hypertension", "DM"))

plot_par <- barplot_diags(df_par, diags_to_remove = c("Arterial hypertension", "DM"))


```


#Vis Diagnosis Occurence (Relative Comparison Control to DOI)
```{r}
# Prepare your data
sum_raw_pos <- left_join(sum_diagnosis, sum_positives, by="Diagnosis") %>%
  filter(Diagnosis != "Hepatobiliary cancer") %>%
  arrange(desc(Percentage.y)) %>%
  mutate(Legend = LETTERS[1:45])

sum_par_pos <- left_join(sum_par, sum_positives, by="Diagnosis") %>%
  filter(Diagnosis != "Hepatobiliary cancer") %>%
  arrange(desc(Percentage.y)) %>%
  mutate(Legend = LETTERS[1:45])


sum_pos <- left_join(sum_diagnosis, sum_positives, by="Diagnosis") %>%
  left_join(sum_par, by="Diagnosis") %>%
  filter(Diagnosis != "Hepatobiliary cancer") %>%
  arrange(desc(Percentage.y)) %>%
  mutate(Legend = LETTERS[1:45])




comparison_plot_diags(sum_pos, base_size=24)

```







#Stacked Bars DOI Etiologies for Liver disease

```{r}
load(paste0(project_path, "/data/dataframes/df_y.RData"))

priority_order <- c("Cirrhosis", "Viral Hepatitis", "CLD", "No Liver disease")
color_map <- c("Cirrhosis" = "#22a884", 
               "Viral Hepatitis" = "#414487", 
               "CLD" = "#fde725", 
               "No Liver disease" = "#b7b7b7")

pat_cld <- pat_icds[pat_icds$diag_icd10 %in% par_icd_codes | pat_icds$diag_icd9 %in% par_icd_codes, ] %>%
  select(c("eid", "diag_icd9", "diag_icd10", "epistart")) %>%
  left_join(Patients_at_risk, by = c("diag_icd10" = "ICD10")) %>%
  right_join(df_y, by = "eid") %>%
  subset(status==1) %>%
  select(-c("location_name", "location_code", "location_nr", "location_country", "country_code", "split_ext", "split_int"))

pat_cld$epistart[is.na(pat_cld$Group)] <- as.Date(pat_cld$date_of_diag)

pat_cld$Group[is.na(pat_cld$Group)] <- "No Liver disease"
pat_cld$Group[!pat_cld$Group %in% par_subset] <- "No Liver disease" #Replace all non-matching groups with "No LD"
pat_cld$Group <- factor(pat_cld$Group, levels=priority_order)

summary(pat_cld$Group)

priority <- function(diagnosis) {
  case_when(
    diagnosis == "Cirrhosis" ~ 1,
    diagnosis == "Viral Hepatitis" ~ 2,
    diagnosis == "CLD" ~ 3,
    diagnosis == "No Liver disease" ~ 4,
    TRUE ~ 5  # Assign a lower priority to other diagnoses
  )
}


pat_cld <- pat_cld %>%
  group_by(eid)

# Node 0 represents first visit to hospital after assessment
pat_cld_node0 <- pat_cld %>%
  mutate(Priority = priority(Group)) %>%
  group_by(eid) %>%
  filter(epistart == min(epistart)) %>%
  arrange(eid, Priority) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(-Priority)

summary_node0 <- pat_cld_node0 %>%
  group_by(Group) %>%
  summarize(Count = n(), .groups = 'drop') %>%
  mutate(Time = "First \nEHR") %>%
  mutate(Order = 1 ) %>%
  mutate(Priority = priority(Group)) %>%
  arrange(Priority) %>%
  mutate(Percentage = round(Count / sum(Count) * 100)) 


#  Node 1 represents summary of all visits (together with prioritising diagnosis, this leaves the most dire diagnosis present)
pat_cld_node1 <- pat_cld %>%
  mutate(Priority = priority(Group)) %>%
  group_by(eid) %>%
  #filter(epistart == max(epistart)) %>%   #better to take all incidents than just the last, as not all diags get coded everytime
  arrange(eid, Priority) %>%
  filter(row_number() == 1) %>%
  ungroup() %>%
  select(-Priority)

summary_node1 <- pat_cld_node1 %>%
  group_by(Group) %>%
  summarize(Count = n(), .groups = 'drop') %>%
  mutate(Time = paste0("Prior to\n", DOI)) %>%
  mutate(Order = 2 ) %>%
  mutate(Priority = priority(Group)) %>%
  arrange(Priority) %>%
  mutate(Percentage = round(Count / sum(Count) * 100))


# View the summaries
print(summary_node0)
print(summary_node1)


# Merge timepoints
combined_data <- rbind(summary_node0, summary_node1)





distribution_plot <- stacked_bars_time_comparison(combined_data, 
                                             base_size = 22, 
                                             priority_order = priority_order,
                                             color_map = color_map)

# Print the plot
print(distribution_plot )

# Save the plot
ggsave(filename = file.path(project_path, paste0("supplement_visuals/", DOI, "_Distributions_diagnosis.svg")), 
       plot = distribution_plot , width = 12, height = 8, bg = "white")
```






```{r}
pat_cld_all <- pat_icds_all[pat_icds_all$diag_icd10 %in% par_icd_codes] %>%
  select(c("eid", "diag_icd10", "epistart")) %>%
  left_join(Patients_at_risk, by = c("diag_icd10" = "ICD10")) 

df_y_only <- df_y[df_y$status==1] %>%
  select(-c("location_name", "location_code", "location_nr", "location_country", "country_code", "split_ext", "split_int"))

pat_cld_all$epistart <- ymd(pat_cld_all$epistart)
df_y_only$date_of_diag <- ymd(df_y_only$date_of_diag)

# Modify pat_cld_all to add date_of_diag as epistart for "No Liver disease" cases
pat_cld_all <- pat_cld_all %>%
  right_join(df_y_only, by = "eid") %>%
  group_by(eid) %>%
  mutate(
    Group = ifelse(all(is.na(Group)), "No Liver disease", first(na.omit(Group))),
    epistart = case_when(
      !is.na(epistart) ~ epistart,
      Group == "No Liver disease" & !is.na(date_of_diag) ~ date_of_diag,
      Group == "No Liver disease" & is.na(date_of_diag) & !is.na(date_of_death) ~ date_of_death,
      TRUE ~ NA_Date_
    )
  ) %>%
  ungroup()

total_patients <- n_distinct(pat_cld_all$eid)
print(paste("Total patients before filtering:", total_patients))



pat_cld_node2 <- pat_cld_all %>%
  filter(epistart <= date_of_diag + years(2))

included_patients <- n_distinct(pat_cld_node2$eid)
print(paste("Patients included after filtering:", included_patients))

missing_patients <- anti_join(pat_cld_all, pat_cld_node2, by = "eid")
print(paste("Number of missing patients:", nrow(missing_patients)))

if (nrow(missing_patients) > 0) {
  print("Details of missing patients:")
  print(missing_patients %>% select(eid, Group, epistart, date_of_diag))
  
  # Add these missing patients back as "No Liver disease"
  missing_patients_added <- missing_patients %>%
    mutate(
      Group = "No Liver disease",
      epistart = date_of_diag  # Set epistart to date_of_diag for these patients
    ) %>%
    select(names(pat_cld_node2))  # Ensure column consistency
  
  pat_cld_node2 <- bind_rows(pat_cld_node2, missing_patients_added)
}



summary_node2 <- pat_cld_node2 %>%
  mutate(Priority = priority(Group)) %>%
  group_by(eid) %>%
  arrange(eid, Priority) %>%
  slice(1) %>%  # Take the first (highest priority) diagnosis for each patient
  ungroup() %>%
  group_by(Group) %>%
  summarize(Count = n(), .groups = 'drop') %>%
  mutate(Time = paste0("0-2 y\nafter ", DOI)) %>%
  mutate(Order = 3) %>%
  mutate(Priority = priority(Group)) %>%
  arrange(Priority) %>%
  mutate(Percentage = round(Count / sum(Count) * 100))


#Diagnostics

print(paste("Total HCC patients:", nrow(df_y_only)))
print(paste("Patients in pat_cld_all:", n_distinct(pat_cld_all$eid)))
print(paste("Patients in pat_cld_node2:", n_distinct(pat_cld_node2$eid)))
print(paste("Patients in summary_node2:", sum(summary_node2$Count)))
missing_patients <- anti_join(pat_cld_all, pat_cld_node2, by = "eid")
print(paste("Number of missing patients:", nrow(missing_patients)))

# Print the summary to check
print(summary_node2)


combined_data <- rbind(combined_data, summary_node2)
updated_plot <- stacked_bars_time_comparison(combined_data, base_size = 22, priority_order = priority_order, color_map = color_map)
ggsave(filename = file.path(project_path, paste0("supplement_visuals/", DOI, "_Distributions_Updated.svg")), 
       plot = updated_plot, width = 12, height = 8, bg = "white")




```


```{r}
# Step 1: Check initial patient count and group distribution
total_patients <- n_distinct(pat_cld_all$eid)

initial_group_distribution <- pat_cld_all %>%
  group_by(eid) %>%
  summarize(
    Groups = list(unique(na.omit(Group))),
    Has_Liver_Disease = length(unique(na.omit(Group))) > 0
  ) %>%
  ungroup() %>%
  mutate(
    Cirrhosis = sapply(Groups, function(x) "Cirrhosis" %in% x),
    Viral_Hepatitis = sapply(Groups, function(x) "Viral Hepatitis" %in% x),
    CLD = sapply(Groups, function(x) "CLD" %in% x),
    No_Liver_Disease = !Has_Liver_Disease
  ) %>%
  summarize(
    Cirrhosis = sum(Cirrhosis),
    `Viral Hepatitis` = sum(Viral_Hepatitis),
    CLD = sum(CLD),
    `No Liver disease` = sum(No_Liver_Disease)
  ) %>%
  pivot_longer(everything(), names_to = "Group", values_to = "n") %>%
  mutate(Percentage = n / total_patients * 100) %>%
  arrange(desc(n))

print("Initial patient count and group distribution:")
print(paste("Total patients:", total_patients))
print(initial_group_distribution)
```


```{r}
pat_cld$epistart[is.na(pat_cld$Group)] <- as.Date(pat_cld$date_of_diag)

pat_cld$Group[is.na(pat_cld$Group)] <- "No Liver disease"
pat_cld$Group[!pat_cld$Group %in% par_subset] <- "No Liver disease" #Replace all non-matching groups with "No LD"
pat_cld$Group <- factor(pat_cld$Group, levels=priority_order)

summary(pat_cld$Group)

priority <- function(diagnosis) {
  case_when(
    diagnosis == "Cirrhosis" ~ 1,
    diagnosis == "Viral Hepatitis" ~ 2,
    diagnosis == "CLD" ~ 3,
    diagnosis == "No Liver disease" ~ 4,
    TRUE ~ 5  # Assign a lower priority to other diagnoses
  )
}
```



























# Ignore stuff below here







Visualization Chronic Liver Disease - HCC
```{r}

# Count unique eids with no entry in diag_icd10
count_removed_eids <- pat_cld %>%
  filter(is.na(diag_icd10)) %>%
  summarise(count = n_distinct(eid)) %>%
  pull(count)

#Calculate for each patient how much time only diagnosis of 

cancer_diag_dates <- pat_cld %>%
  select(eid, date_of_diag) %>%
  distinct()  # Assuming one cancer diagnosis date per patient


pat_cld <- pat_cld %>%
  filter(!is.na(diag_icd10)) %>%
  select(-date_of_diag) %>%
  left_join(cancer_diag_dates, by = "eid") %>%
  mutate(
    epistart = as.Date(epistart, format = "%d/%m/%Y"),
    date_of_diag = as.Date(date_of_diag, format = "%Y-%m-%d"),
  ) %>%
  arrange(eid, epistart)

pat_cld <- pat_cld %>%
  group_by(eid) %>%
  arrange(eid, epistart) %>%
   mutate(
    next_eid = lead(eid),
    next_epistart = lead(as.Date(epistart, format = "%d/%m/%Y")),
    end_date = dplyr::if_else(
      eid == next_eid & epistart < date_of_diag, 
      next_epistart, 
      date_of_diag, 
      missing = as.Date(NA)
    ),
    duration = ifelse(is.na(end_date) & epistart < date_of_diag, as.numeric(date_of_diag - epistart),
                      ifelse(epistart < date_of_diag, as.numeric(end_date - epistart), 0))
  ) #%>%
  #ungroup() #%>%
 # select(-next_eid, -next_epistart)
```




















```{r}

library(networkD3)
library(dplyr)
 
# A connection data frame is a list of flows with intensity for each flow
links <- data.frame(
  source=c("group_A","group_A", "group_B", "group_C", "group_C", "group_E"), 
  target=c("group_A","group_D", "group_E", "group_F", "group_G", "group_H"), 
  value=c(2,3, 2, 3, 1, 3)
  )
 
# From these flows we need to create a node data frame: it lists every entities involved in the flow
nodes <- data.frame(
  name=c(as.character(links$source), 
  as.character(links$target)) %>% unique()
)
 
# With networkD3, connection must be provided using id, not using real name like in the links dataframe.. So we need to reformat it.
links$IDsource <- match(links$source, nodes$name)-1 
links$IDtarget <- match(links$target, nodes$name)-1
 
# Make the Network
p <- sankeyNetwork(Links = links, Nodes = nodes,
              Source = "IDsource", Target = "IDtarget",
              Value = "value", NodeID = "name", 
              sinksRight=FALSE)
p


```






```{r}
plot_pie_chart <- function(data, title = "Pie Chart") {
  ggplot(data, aes(x = "", y = Count, fill = Group)) +
    geom_bar(width = 1, stat = "identity") +
    coord_polar("y", start = 0) +
    geom_text(aes(label = Count), position = position_stack(vjust = 0.5),
              size = 12,  # Increase the text size for numbers
              color = "black") +
    scale_fill_brewer(palette = "Set3") +  # Using a Brewer color palette
    theme_void() +
    theme(legend.title = element_blank(),
          legend.text = element_text(size = 30),  # Increase the text size for legend
          legend.position = "right") +  # Adjust legend position
    labs(title = title, fill = "Group")
}

# Example usage with your tibble
pie_chart_node0 <- plot_pie_chart(summary_node0, title = "Node 0 Distribution")
print(pie_chart_node0)
setwd(project_path)
ggsave(filename = file.path(project_path, paste0("supplement_visuals/", DOI, "_Distributions_0.svg")), 
       plot = pie_chart_node0, width = 10, height = 7, bg = "white")


pie_chart_node1 <- plot_pie_chart(summary_node1, title = "Node 1 Distribution")

# To display the plots

print(pie_chart_node1)
ggsave(filename = file.path(project_path, paste0("supplement_visuals/", DOI, "_Distributions_1.svg")), 
       plot = pie_chart_node1, width = 10, height = 7, bg = "white")

```













































# Ignore this for now








```{r}
library(dplyr)

pat_cld_summary <- pat_cld_summary %>%
  # Join with the total durations data
  left_join(total_durations, by = "eid") %>%
  # Add a rank column based on conditions
  group_by(eid) %>%
  mutate(
    has_viral_hepatitis = "Viral Hepatitis" %in% Etiology,
    has_cirrhosis = "Cirrhosis" %in% Etiology,
    has_sld = "SLD" %in% Etiology,
    rank = case_when(
      has_viral_hepatitis & !has_cirrhosis & !has_sld & total_duration_eid > 1 ~ 1,
      has_viral_hepatitis & has_cirrhosis & total_duration_eid > 1 ~ 2,
      has_cirrhosis & !has_viral_hepatitis & !has_sld ~ 3,
      has_sld & has_cirrhosis ~ 4,
      has_sld & !has_cirrhosis & !has_viral_hepatitis ~ 5,
      TRUE ~ 6  # Default rank if none of the above conditions are met
    )
  ) %>%
  ungroup() %>%
  arrange(rank, desc(total_duration_eid))

plot <- ggplot(pat_cld_summary, aes(x = reorder(factor(eid), rank), y = as.numeric(total_duration), fill = Etiology)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "Patient ID", y = "Duration (Days)", fill = "Condition") +
  theme(
    legend.position = "bottom",
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()
  )

print(plot)






# Calculate total duration for each eid
total_durations <- pat_cld_summary %>%
  group_by(eid) %>%
  summarise(total_duration_eid = sum(as.numeric(total_duration))) %>%
  arrange(desc(total_duration_eid))

# Reorder the eids based on total duration
pat_cld_summary$eid <- factor(pat_cld_summary$eid, levels = total_durations$eid)


# Create the sorted stacked bar plot
plot <- ggplot(pat_cld_summary, aes(x = eid, y = as.numeric(total_duration), fill = Etiology)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  labs(x = "Patient ID", y = paste("Duration (Days) until" , DOI), fill = "Condition") +
  theme(
    legend.position = "bottom",  # Adjust legend position
    axis.text.x = element_blank(),  # Remove x-axis labels
    axis.ticks.x = element_blank(),  # Optionally remove x-axis ticks
    #axis.title.x = element_blank()   # Remove x-axis title
  )

print(plot)

# Save the plots
ggsave(filename = file.path(supplement_visuals_dir, paste0("ICD_Intervall_until_", IOI, "_", Sys.Date(), ".png")), 
       plot = plot, width = 12, height = 8, bg= "white")
ggsave(filename = file.path(supplement_visuals_dir, paste0("ICD_Relations_", IOI, "_", Sys.Date(), ".svg")), 
       plot = plot, width = 12, height = 8, bg= "white")

# pat_cld_wide <- pat_cld_summary %>%
#   pivot_wider(
#     names_from = Etiology,
#     values_from = total_duration,
#     values_fill = list(total_duration = as.difftime(0, units = "days"))  # Fill missing values with 0 days
#   )
```


```{r}
pat_cld <- pat_icds[pat_icds$diag_icd10 %in% par_icd_codes | pat_icds$diag_icd9 %in% par_icd_codes, ] %>%
  select(c("eid", "diag_icd9", "diag_icd10", "epistart")) %>%
  left_join(ICD_Singles, by = c("diag_icd10" = "ICD10")) %>%
  right_join(df_y, by = "eid")


summary_table <- pat_cld %>%
  group_by(diag_icd10) %>%
  summarize(Count = n()) %>%
  left_join(ICD_Singles, by = c("diag_icd10" = "ICD10")) %>%
  select(diag_icd10, Count, Diagnosis)

pat_cld <- left_join(df_y, pat_cld) %>%
  mutate(epistart = as.Date(epistart, format = "%d/%m/%Y")) %>%
  group_by(eid) %>%
  slice_min(epistart, n = 1)


pat_cld$date_of_diag <- as.Date(pat_cld$date_of_diag)
pat_cld$duration <- as.numeric(difftime(pat_cld$date_of_diag, pat_cld$epistart, units = "days"))
pat_cld <- pat_cld[order(pat_cld$duration), ]
pat_cld$eid <- with(pat_cld, reorder(as.factor(eid), -duration))
pat_cld$duration_years <- pat_cld$duration / 365.25

pat_cld <- pat_cld %>%
  arrange(Group, duration_years)

# Plot
plot <- ggplot(pat_cld, aes(x = reorder(as.factor(eid), duration_years), y = duration_years, fill = Group)) +
  geom_bar(stat = "identity", position = "dodge") +  # Use dodge to separate the bars by groups
  labs(
    title = "Duration between diagnosis of Chronic Liver Disease and HCC in Years",
    x = "",
    y = "Duration (Years)"
  ) +
  scale_fill_discrete(name = "Group") +  # Add a legend for the groups
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),  # Remove x-axis labels
    axis.ticks.x = element_blank(),  # Optionally remove x-axis ticks
    axis.title.x = element_blank()   # Remove x-axis title
  )
plot

# Create the 'supplement_visuals' directory
supplement_visuals_dir <- file.path(project, "supplement_visuals")
dir.create(supplement_visuals_dir, showWarnings = FALSE) # Create the directory if it doesn't already exist

# Save the plots
ggsave(filename = file.path(supplement_visuals_dir, paste0("Timespan from CLD to ", DOI, ".png")), 
       plot = plot, width = 10, height = 12, bg= "white")
ggsave(filename = file.path(supplement_visuals_dir, paste0("Timespan from CLD to ", DOI, ".svg")), 
       plot = plot, width = 10, height = 12, bg = "white")

plot

```


```{r}
# Plot
plot <- ggplot(pat_cld, aes(x = epistart, xend = date_of_diag, y = as.factor(eid), yend = as.factor(eid))) +
  geom_segment(aes(color = duration < 0), size = 0.3) +  # Color lines with negative duration in red
  scale_color_manual(values = c("black", "red")) +  # Set custom colors
  labs(
    title = "Timespan between diagnosis of Chronic Liver Disease and HCC",
    x = "Time",
    y = "EID"
  ) +
  theme_minimal() +
  theme(axis.text.y = element_blank(),  # Remove y-axis tick labels
        legend.position = "none")  # Remove legend

# Create the 'supplement_visuals' directory
supplement_visuals_dir <- file.path(project, "supplement_visuals")
dir.create(supplement_visuals_dir, showWarnings = FALSE) # Create the directory if it doesn't already exist

# Save the plots
ggsave(filename = file.path(supplement_visuals_dir, paste0("Timespan from CLD to ", DOI, ".png")), 
       plot = plot, width = 10, height = 12, bg= "white")
ggsave(filename = file.path(supplement_visuals_dir, paste0("Timespan from CLD to ", DOI, ".svg")), 
       plot = plot, width = 10, height = 12, bg = "white")

plot
```



