```{r}
library(MatchIt)
library(dplyr)
library(ggplot2)
library(stats)
library(base)
library(broom)
library(tidyr)
library(readr)
library(dplyr)
library(gtools)
library(data.table)
options(java.parameters = "-Xmx8000m")
library(rJava)
library(xlsx)
library(readxl)
if (!requireNamespace("MatchIt", quietly = TRUE)) {
  install.packages("MatchIt")
}
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}

library(MatchIt)
library(dplyr)
source("../config.R") #Loads libraries, variables and global function
```


#### Load required Dataframes

```{r}
setwd(sharepoint_ukb)
proteomics_eids <- read.csv("C:/Users/felix/OneDrive - Uniklinik RWTH Aachen/public/ukb/extracted/df_proteomics_eids.csv")
cols_to_read <- c("eid")
#proteomics_eids <- fread("raw/olink_data_full.txt", select = cols_to_read) %>% #check_and_remove_withdrawals(df_withdrawals)
# proteomics_eids <- proteomics_eids %>% filter(!duplicated(eid))
# write.csv(proteomics_eids, file = "C:/Users/felix/OneDrive - Uniklinik RWTH Aachen/public/ukb/extracted/df_proteomics_eids.csv", col.names=TRUE, row.names=TRUE, append=FALSE)
setwd(project_path)
load("data/dataframes/df_covariates.RData")
load("data/dataframes/df_y.RData")

```


#### Build Propensity Score Dataframe with your wanted covariates and Status variable
```{r}
df_psm <- df_covariates %>% select(c(eid, AGE, SEX, BMI, MultipleDeprivationIndex))
df_status <- df_y %>% select(c("eid", "status"))
df_psm <- merge(df_psm, df_status, by = "eid")
#df_psm <- merge(proteomics_eids, df_psm, by="eid")
df_psm <- df_psm %>% mutate(SEX = ifelse(SEX == "Male", 1, ifelse(SEX == "Female", 0, NA)))
sum(df_psm$status)
df_psm <- na.omit(df_psm) #only works with no NANs in your df
```


#### Matching your cases, you can adjust the ratio and your covariates that you want to include   #### here
```{r}

propensity_model <- glm(status ~ AGE + SEX + BMI + MultipleDeprivationIndex, data = df_psm, family = binomial())
df_psm$propensity_score <- predict(propensity_model, type = "response")


cca_match<-matchit(status~  AGE + SEX + BMI+MultipleDeprivationIndex, data=df_psm, method="nearest", ratio=2, caliper=0.2)
cca_match10<-matchit(status~  AGE + SEX + BMI+MultipleDeprivationIndex, data=df_psm, method="nearest", ratio=10, caliper=0.2)
cca_match20<-matchit(status~  AGE + SEX + BMI+MultipleDeprivationIndex, data=df_psm, method="nearest", ratio=20, caliper=0.2)

cca_gematchedAgeSexBMIMDI <- match.data(cca_match)
cca_gematched10AgeSexBMIMDI <- match.data(cca_match10)
cca_gematched20AgeSexBMIMDI <- match.data(cca_match20)
```


#### Check your Dataframe and save it as a csv for further analyses

```{r}
df_psm
```

```{r}
write.csv(cca_gematched10AgeSexBMIMDI, file="C:/Users/felix/OneDrive - Uniklinik RWTH Aachen/public/projects/cca/data/dataframes/df_psm_cca10.csv", col.names=TRUE, row.names=TRUE, append=FALSE)



##### Select only EID and Status column for the use of Niharikas Tool

cca_niharika_matched <- cca_gematched10AgeSexBMIMDI %>% select(c("eid", "status"))

cca_niharika_matched$status<- as.integer(cca_niharika_matched$status)
typeof(cca_niharika_matched$status)

write.csv(cca_niharika_matched, file="C:/Users/felix/OneDrive - Uniklinik RWTH Aachen/public/projects/cca/data/dataframes/df_psm_cca10_niharika.csv", col.names=TRUE, row.names=TRUE, append=FALSE)



splenomegaly <-read.csv("C:/Users/felix/OneDrive - Uniklinik RWTH Aachen/Niharika Tool Proteomics/volcano_plot_generator/labels/Splenomegaly.csv")
typeof(splenomegaly$)
```



# Eigentliches Matching (change ratio for amount of matches for each HCC)
HCC_match20<-matchit(HCC~  AGE + SEX + BMI, data=Pat_with_metabolomics, method="nearest", ratio=20, caliper=0.2)
HCC_gematched20AgeSexBMI <- match.data(HCC_match20)

