---
title: "Table Y"
output: html_notebook
---

```{r}
source("../config.R") #Loads libraries, variables and global function
```

#### Load hesin data (NHS Information on hospital admissions, diagnosis etc)
```{r}
#lade die Datensets mit Diagnosen der Patienten (more info here: https://biobank.ndph.ox.ac.uk/showcase/refer.cgi?id=141140 )
setwd(sharepoint_ukb)
df_withdrawals <- read.delim(withdrawals_path)
cols_to_read <- c("eid", "ins_index", "diag_icd10", "diag_icd9") #predefine cols for hesin/hesin_diag to be read
hesin_diag <- fread(hesin_diag_path, select = cols_to_read) %>% check_and_remove_withdrawals(df_withdrawals)

cols_to_read <- c("eid", "ins_index", "epistart")
hesin <- fread(hesin_path, select = cols_to_read) %>% check_and_remove_withdrawals(df_withdrawals)

#merge hesin tables (unique identifier for every patient admission by eid + ins_index)
setkey(hesin_diag, eid, ins_index)
setkey(hesin, eid, ins_index)
pat_icds <- merge(hesin_diag, hesin, by = c("eid", "ins_index"), all = TRUE)

#Access file on death causes (reported independently to NHS -> UKB, so there might be additional patients with your IOI in this dataset)
setwd(sharepoint_ukb)
death_cause <- read.delim(death_cause_path) %>%
  select(eid, cause_icd10)
death_date <- read.delim(death_date_path) %>% 
  select(eid, date_of_death)

death_info <- merge(death_cause, death_date, by="eid") %>% # Merge the data frames
  check_and_remove_withdrawals(df_withdrawals)
rm(hesin, hesin_diag, death_cause, death_date) #Clean up tables no longer in use
sub_icd <- subset(pat_icds, diag_icd10 %in% IOIs)

#Export sub_icd if running through this multiple times because the part above takes quite some time due to large dfs
setwd(sharepoint_ukb)
filename <- paste("processed/sub_icd_" ,DOI, ".csv")
#write.csv(sub_icd, file = filename, row.names = FALSE) 
```


#### Define function subset_hesin to Extract information on DOI, add data on "died from DOI" to "was diagnosed with DOI"
```{r}

summary_df <- data.frame(icd_code = character(), amount = numeric())

setwd(sharepoint_ukb)
subset_hesin <- function(IOI, DOI, summary_df) {
      #Subset for only rows in hesin with icd-code of interest (IOI)
      sub_icd <- subset(pat_icds, diag_icd10 == IOI)
      print(paste("Amount of ", IOI, " in UKB w/o death info: ",length(unique(sub_icd$eid))))
      sub_icd$epistart <- as.Date(sub_icd$epistart, format="%d/%m/%Y") #transform to date       object
      sub_icd$year <- format(sub_icd$epistart, "%Y") #extract year
      sub_icd$year <- as.numeric(format(sub_icd$epistart, "%Y"))
      
        # the arrange() function sorts the rows in ascending order  based on eid, diag_icd10, and epistart. The is.na(epistart) term ensures that NAs will be sorted to the end. Finally, slice_head(n = 1) picks the first row for each eid and diag_icd10 combination, which will be the row with the earliest epistart or year value.
      
      sub_icd <- sub_icd %>%
        group_by(eid, diag_icd10) %>%
        arrange(eid, diag_icd10, is.na(epistart), epistart) %>%
        slice_head(n = 1) %>%
        ungroup() %>%
        select(eid, diag_icd10, epistart, year)
      
      sub_death <- subset (death_info, cause_icd10 == IOI) #subset death_table
      sub_death$date_of_death <- as.Date(sub_death$date_of_death, format="%d/%m/%Y")
      
      #For those without diagnosis but only death date, use death date (dont use ifelse because date-attribute will be stripped)
      sub <- merge(sub_icd, sub_death , by.x = "eid", by.y = "eid", all.x=TRUE, all.y=TRUE) %>%
        rename("date_of_diag" = "epistart") %>%
        mutate(date_of_diag = case_when(is.na(date_of_diag) ~ date_of_death, TRUE ~ date_of_diag))
      
      sub <- as.data.frame(sub[!duplicated(sub$eid),]) %>%      # Remove duplicates
        mutate(diag_icd10 = coalesce(diag_icd10, cause_icd10))  # enter diag info when only "died from info" is available
        new_row <- data.frame(icd_code = IOI, amount = nrow(sub))
  summary_df <<- rbind(summary_df, new_row) #feed summary_df

      colnames(sub)[2]<- "status"
      sub[,2] <- 1                                  #Patienten mit Diagnose X erhalten Wert 1                                                              
      print(paste("Amount of ", IOI, " in UKB: ", count(sub)))
      print(paste("Amount of ", IOI, " in Death Register: ", count(sub_death)))
      count <- count(sub)
      sapply(sub, class)
      
  return(sub) # Return the sub and sub_grouped dataframes
}


#Execute the function for all IOIs
df_list <- list()
for(IOI in IOIs) {
  df_name <- paste0("sub_", IOI) # Dynamically create a variable name
  new_df <- subset_hesin(IOI, DOI, summary_df) # Call the function and assign the result to a new dataframe
  assign(df_name, new_df)
  df_list[[df_name]] <- new_df
}

# Combine all dataframes into a single dataframe "sub"
sub <- do.call(rbind, df_list)

# Calculate the number of duplicated eids in sub
count_eid_duplications(sub, df_name= "Subset of Patients with DOI")

# Check later on this!!!! ##############
# print(paste("Number of duplicated eids:", duplicated_eids_count))
# summary_df <- rbind(summary_df, data.frame(icd_code = "duplicated", amount = duplicated_eids_count)) #store duplicated info in summary
# 
# df_discard <- df_discard%>% distinct(eid, .keep_all = TRUE)




```

#### Access all eids as well as info on assessment center + date of first visit
```{r}
#Access all UKB_eids either from sharepoint
  #setwd(sharepoint_ukb)
  #df <- fread("raw/ukb52200.csv", select=c("eid", "52-0.0", "34-0.0" "53-0.0", "54-0.0"))   
  #write.csv(df, file="extracted/basic_info.csv", row.names=FALSE)   
  
#shortcut
setwd(sharepoint_ukb)
df_basic <- fread("extracted/basic_info.csv") %>% check_and_remove_withdrawals(df_withdrawals)


# setwd(sharepoint_ukb)
# write.csv(df_basic, "extracted/basic_info.csv", row.names=FALSE)
  
#access detailled info on different ukb_assessment centers 
location <- read_excel(master_table, sheet = "UKB_Center") %>%
  filter(!grepl("imaging", location_name, ignore.case = TRUE))

try({
  df_basic <- df_basic %>%
    rename_with(
      ~ case_when(
        . == "53-0.0" ~ "assessment",
        . == "54-0.0" ~ "location_code",
        . == "52-0.0" ~ "month_of_birth",
        . == "34-0.0" ~ "year_of_birth",
        TRUE ~ .
      )
    ) %>%
    mutate(across(assessment, ~as.Date(., format = "%Y-%m-%d")))
  
  print("Data cleaning completed successfully.")
})

# Merge df_basic with sub (which is "raw" df_y) 
  df_discard <- merge(df_basic, sub, by.x = "eid", by.y = "eid", all.x=TRUE)
  sapply(df_discard, class)
  
# Label all w/o DOI as 0, those with DOI 1 year after first visit as 1, rest as 2 (these will have to be removed, also in df_X)
  # Replace NA's with 0 in 'status'
df_discard$status[is.na(df_discard$status)] <- 0

#Create a column that calculates the difference in time between assessment and diagnosis date (all cases with positive difftime (or negative but > -1 have to be discarded))
df_discard <- df_discard %>%
  mutate(difftime = ifelse(status == 1, difftime(assessment, date_of_diag, units = "weeks") / 52.25, NA))

#label 2 for all DOI diagnosis previous to UKB assessment
df_discard <- df_discard %>%
  mutate(status = case_when(
    status == 1 & (difftime > 0 | (difftime >= -1 & difftime < 0)) ~ 2, TRUE ~ status),
  discard = ifelse(status == 2, 1, 0))
df_discard <- df_discard%>% distinct(eid, .keep_all = TRUE)
#save(df_discard, file=paste(project_path, "/data/dataframes/df_discard.RData", sep=''))
```

#### Prepare df_y
```{r}
setwd(project_path)
if (!exists("df_discard")) load("data/dataframes/df_discard.RData") #loads df_discard from project directory (to skip parts above)

# Count and print the number of discarded cases
doi_discarded <- sum(df_discard$discard)
print(paste("Number of cases of", DOI, "discarded because of diagnosis before/during assessment time:", doi_discarded))

df_y <- subset(df_discard, status != 2)
doi_included <- sum(df_y$status)
print(paste("Number of cases of", DOI, "still available after removing previously diagnosed people:", doi_included))
df_y_dupes <- df_y[duplicated(df_y$eid)]
df_y <- df_y[!duplicated(df_y$eid)]
doi_included <- sum(df_y$status)
print(paste("Number of", DOI, "still included after removing duplicates:", doi_included))

mean_difftime <- round((mean(df_y$difftime[df_y$status == 1], na.rm = TRUE) * -1), 2)
print(paste("Average time from assessment to", DOI, ":", mean_difftime, "years"))

```
#### Visual AGE of HCC Patients

```{r}
plot_hcc_diagnosis_age <- function(df, n_total, base_size = 18) {
  # Calculate age at diagnosis or death
  df <- df %>%
    mutate(
      year_of_birth = as.integer(year_of_birth),
      month_of_birth = as.integer(month_of_birth),
      birth_date = ymd(paste(year_of_birth, month_of_birth, "01")),
      diagnosis_date = if_else(!is.na(date_of_diag), date_of_diag, date_of_death),
      age_at_diagnosis = interval(birth_date, diagnosis_date) / years(1)
    ) %>%
    filter(!is.na(age_at_diagnosis) & age_at_diagnosis >= 0)
  
  age_summary <- summary(df$age_at_diagnosis)
  age_sd <- sd(df$age_at_diagnosis)
  cat("Summary statistics for age at HCC diagnosis:\n")
  cat("Min:", round(age_summary["Min."], 2), "\n")
  cat("Max:", round(age_summary["Max."], 2), "\n")
  cat("Mean:", round(age_summary["Mean"], 2), "\n")
  cat("SD:", round(age_sd, 2), "\n")

  # Calculate min and max ages, rounded to nearest 5
  min_age <- floor(min(df$age_at_diagnosis) / 5) * 5
  max_age <- ceiling(max(df$age_at_diagnosis) / 5) * 5

  # Create the plot
  plot <- ggplot(df, aes(x = age_at_diagnosis)) +
    geom_histogram(binwidth = 1, color = "black", fill = "grey", size = 0.2) +
    xlab("Age") +
    ylab("Absolute number of cases") +
    ggtitle("Age at first HCC Diagnosis (UKB)") +
    theme_minimal() +
    theme(
      plot.title = element_text(hjust = 0.5, size = base_size),
      axis.title.x = element_text(size = base_size, vjust=-1),
      axis.title.y = element_text(size = base_size, vjust=1.2),
      axis.title.y.right = element_text(size = base_size, angle=90, vjust=-0.5),
      axis.text.x = element_text(size = base_size, colour = "black", vjust = 0.1),
      axis.text.y = element_text(size = base_size, colour = "black"),
      axis.text.y.right = element_text(size = base_size, colour = "black"),
      panel.grid.major = element_blank(),
      panel.grid.minor = element_blank(),
      plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
      panel.border = element_rect(colour = "black", fill = NA, linewidth = 1)
    ) +
    scale_x_continuous(
      limits = c(min_age, max_age),
      breaks = seq(min_age, max_age, by = 5),
      minor_breaks = NULL
    ) +
    scale_y_continuous(
      expand = c(0, 0), 
      limits = c(0, NA),
      sec.axis = sec_axis(
        ~. / n_total * 100000,
        name = "Incidence [n / 100.000]"
      )
    )
    
  # Print the plot
  print(plot)
  
  # Save the plot
  ggsave(filename = file.path(supplement_visuals_dir, paste0(DOI, "_diagnosis_age_histogram.svg")), 
         plot = plot, width = 10, height = 10, bg = "white")
  
  return(plot)
}

# Usage
df_age_hcc <- subset(df_y, status == 1)

n_total <- nrow(df_y)
plot_hcc_diagnosis_age(df_age_hcc, n_total=n_total, base_size = 30)
```




#### Visualize time of HCC diagnosis and whether patient included in model or not

```{r}

plot_included_discarded_cases(df_discard, base_size=27)

```






#### Preprocess locations
```{r}
# Load the 'world.cities' data
data("world.cities")

# Create a mapping of city names to latitudes and longitudes
city_locations <- world.cities %>%
  filter(country.etc == "UK") %>%
  select(name, lat, long) %>%
  rename(location_name = name)

#add manual data for croydon/newcastle, because they miss in the maps dataset
city_locations <- city_locations %>%
   rbind(data.frame(location_name = "Barts", lat = 51.5177, long = -0.0968),  # Barts (angenommen, es handelt sich um St Bartholomew's Hospital in London) 
        data.frame(location_name = "Hounslow", lat = 51.4664, long = -0.3509),  # Hounslow's coordinates
        data.frame(location_name = "Stoke", lat = 53.0027, long = -2.1799),  # Stoke's coordinates
        data.frame(location_name = "Croydon", lat = 51.3762, long = -0.0982), # Croydon's coordinates
        data.frame(location_name = "Newcastle", lat = 54.9784, long = -1.6174)) # Newcastle's coordinates

# Merge a subset of df_y with the location mapping
df_loc <- df_y[df_y$status==1]
df_loc <- merge(df_loc, location, by = "location_code", all = TRUE)
df_loc <- df_loc[!is.na(df_loc$eid)]
df_loc <- df_loc[!duplicated(df_loc$eid)&!duplicated(df_loc$eid),]
len <- nrow(df_loc)

#summarize per country
df_country <- df_loc %>%
    group_by(location_country) %>%
    summarise(count = n()) %>%
    mutate(percentage = (count / len) * 100)

# Count the number of cases in each location
df_loc_counts <- df_loc %>%
  group_by(location_name) %>%
  summarise(count = n())

# Merge the df_loc_counts with the city locations
df_loc_counts <- merge(df_loc_counts, city_locations, by = "location_name", all.x = TRUE)
df_loc_counts <- merge(df_loc_counts, location, by = "location_name", all.x = TRUE)
```




# Quality check distribution over locations and create columns for internal and external group splits
-> External split: Choose ~ 20 % to leave aside while training/hyperparameter tuning for validation in 2nd step. Ideally data from separate countries
```{r}
#Define "external validation centers"
validation <- c("Glasgow", "Edinburgh", "Cardiff", "Swansea", "Wrexham", "Newcastle")

# Merge on location info
df_y <- merge(df_y, location, by = "location_code", all.x = TRUE)

######## Outer level ############
df_y$split_ext <- ifelse(df_y$location_name %in% validation, 1, 0)


######### Inner level ##########
# Sort the centers by amount of DOI to achieve even numbers in every fold
df_loc_train <- df_loc_counts %>%
  filter(!location_name %in% validation) %>%
  arrange(desc(count))

num_folds <- 5
df_loc_train$split_int <- rep(1:num_folds, length.out=nrow(df_loc_train))

sum(df_loc_counts$count) 

# Check "summary_by_fold for even distribution
summary_by_fold <- df_loc_train %>%
  group_by(split_int) %>%
  summarise(centers = paste(location_name, collapse = ", "), 
            total_count = sum(count))
print(sum(summary_by_fold$total_count))

summary_by_fold


#Add split_int info to df_y
df_y <- df_y %>%
  left_join(select(df_loc_train, location_code, split_int), by = "location_code")


#Check absolute and relative amounts
summary_by_level <- df_y %>%
  filter(status == 1) %>%
  group_by(split_ext) %>%
  summarize(
    count = n(),
    relative = n() / nrow(filter(df_y, status == 1))
  )

summary_by_level

split_int <- summary_by_level %>%
  filter(split_ext == 0) %>%
  pull(count)

split_ext <- summary_by_level %>%
  filter(split_ext == 1) %>%
  pull(count)

df_loc_counts$in_validation <- df_loc_counts$location_name %in% validation
df_loc_counts$rank <- with(df_loc_counts, order(in_validation, -count))
df_loc_counts$location_name <- factor(df_loc_counts$location_name, levels = df_loc_counts$location_name[order(df_loc_counts$rank)])

validation

summary_df <- df_y %>% #Group by center and count
  group_by(location_name) %>%
  summarise(
    all = n(),  # Count of all participants in each location
    cases = sum(status == 1, na.rm = TRUE),  # Count of cases where status == 1
    ratio = cases / all *100  # Ratio of cases to total
  )




```

# Visualize locations
```{r}
create_map_plot(df_loc_counts, df_country, base_size=14) #See visual in preprocessing_functions
custom_labels <- paste0(unique(df_country$location_country), " (n=", df_country$count, ")")


# Make the bar plot
plot <- ggplot(df_loc_counts, aes(x = location_name, y = count, fill = location_country)) +
  geom_col() +
  scale_fill_manual(values = c("England" = "grey", "Scotland" = "red", "Wales" = "blue"), labels = custom_labels) +
  coord_flip() +
  theme_minimal() +
  labs(title = paste("Number of ", DOI, " Diagnosis in UK Biobank Centers"),
       x = "Location",
       y = "Number of Cases",
       fill = "Country")

print(plot)



# Save the plots
ggsave(filename = file.path(supplement_visuals_dir, paste0("barplot_", DOI, "_occurence_per_center.png")), 
       plot = plot, width = 10, height = 8, bg= "white")
ggsave(filename = file.path(supplement_visuals_dir, paste0("barplot_", DOI, "_occurence_per_center.svg")), 
       plot = plot, width = 10, height = 7, bg = "white")


```



### Check cancer registry for overlap with ICD10 coded diagnosis

```{r}
setwd(sharepoint_ukb)
df_cancerreg <- fread("raw/ukb677285.csv") %>% check_and_remove_withdrawals(df_withdrawals)
df_y_only <- read.csv(file.path(project_path, "data", "dataframes", "df_y.csv")) %>%
  select("eid", "status") %>%
  subset(status== "1")

df_cancerreg_icd10 <- df_cancerreg %>% select(eid, starts_with("40006")) #Subset for relevant columns for matching with icd10 codes


#Extract for every IOI in IOIs the info on occurence in cancer register
results_list <- list() #Store eids of matched rows in list
for(IOI in IOIs) {
  matched_eids <- df_cancerreg_icd10 %>%                    # Filter rows where any of the columns contain the IOI
    filter(apply(., 1, function(row) any(row == IOI))) %>%
    pull(eid) 
  results_list[[paste(IOI, "cancerreg", sep = "_")]] <- matched_eids
  print(paste("Number of entries for", IOI, ":", length(matched_eids)))
}

matches_count <- numeric(length(IOIs))
names(matches_count) <- IOIs # Naming the vector elements after IOIs for clarity




# out of the eids that are extracted above, create a DF with 0 or 1 per IOI

df_y_cancerreg <- data.frame(eid = df_basic$eid)
for (i in seq_along(results_list)) {
  # Get the current vector of eids
  current_eids <- results_list[[i]]
  column_name <- strsplit(names(results_list)[i], "_")[[1]][1]
  
  # Create the new column with 1 if 'eid' is in the current vector, otherwise 0
  df_y_cancerreg[[column_name]] <- ifelse(df_y_cancerreg$eid %in% current_eids, 1, 0)
}

# Merge IOIs together (if any = true -> status = true)
df_y_cancerreg$status_cancerreg <- apply(df_y_cancerreg[IOIs], 1, function(row) {
  if (any(row == 1)) {
    return(1)
  } else {
    return(0)
  }
})

df_y_cancerreg <- df_y_cancerreg %>%
  select("eid", "status_cancerreg")

df_y <- inner_join(df_y, df_y_cancerreg, by = "eid")

```




#### Export all
```{r}
setwd(project_path)

write.csv(df_y, file=paste(project_path, "/data/dataframes/df_y.csv", sep=''), row.names=FALSE)
save(df_y, file=paste(project_path, "/data/dataframes/df_y.RData", sep='')) 
write.csv(df_loc_counts, file=paste(project_path, "/supplement/", DOI, "_per_center.csv", sep=''), row.names=FALSE)
write.xlsx(summary_df, file=paste0(project_path, "/supplement/", DOI ,"_per_center.xlsx"))
```



































```{r}
# Count matches with other df_y

for(IOI in IOIs) {
  # Extract the vector of eids for the current IOI from the results list
  current_eids <- results_list[[paste(IOI, "cancerreg", sep = "_")]]
  
  # Count how many of these eids are present in df_y_only
  matches_count[IOI] <- sum(df_y_only$eid %in% current_eids)
}

# Print the counts
print(matches_count)






rows_with_C22 <- apply(df_cancerreg_icd10, 1, function(row) {
  any(grepl("^C22", row))
})

# Extract 'eid's from these rows
eids_with_C22 <- df_cancerreg_icd10$eid[rows_with_C22]

# Step 3: Check which of these 'eid's are also in 'df_y_only'
matching_eids <- eids_with_C22[eids_with_C22 %in% df_y_only$eid]

# Step 4: Report the number of matching 'eid's
print(paste("Number of matching 'eid's found:", length(matching_eids)))

#write.csv(data.frame(current_eids), file.path(project, "data/dataframes/cancerreg_single.csv"))
writeLines(as.character(current_eids), con = file.path(project_path, "data/dataframes/cancerreg_single.txt"))

print(data.frame(current_eids))

```







```{r}
setwd(project_path)
load("data/dataframes/df_covariates.RData")
df_covariates <- check_and_remove_withdrawals(df_covariates, df_withdrawals)
df_covariates <- df_covariates %>% select(c(eid, SEX))

sub_icd <- merge.data.frame(sub_icd, df_covariates, by = "eid")
PCD <- as.data.frame(table(sub_icd$diag_icd10))
names(PCD) <- c("Diagnosis", "Freq")
piechart <- ggplot(PCD, aes(x = "", y = Freq, fill = Diagnosis)) +
  geom_bar(width = 1, stat = "identity") +
  coord_polar(theta = "y") +
  theme_void() +
  scale_fill_brewer(palette = "Pastel1") +  # Farbschema
  ggtitle("ICD-10 Diagnosis Distribution") +
  theme(plot.title = element_text(hjust = 0.5))

ggsave(filename = file.path(supplement_visuals_dir, paste0("piechart_", DOI, "_diagnosis_distribution.png")), 
       plot = piechart, width = 10, height = 8, bg= "white")
ggsave(filename = file.path(supplement_visuals_dir, paste0("piechart_", DOI, "_diagnosis_distribution.svg")), 
       plot = piechart, width = 10, height = 7, bg = "white")
```



